On centrality of even orthogonal $\Kt$

### 2016-04-27 paper.tex
Fixed page 7

\begin{rk} \label{rk:T123'} Notice that from the proof of the above proposition it follows that one could replace the family of relations \eqref{add3}
in the definition of $\St(n, R, I)$ with the following smaller family \eqref{add3'}.
\setcounter{equation}{2} \renewcommand{\theequation}{T\arabic{equation}'}
\begin{equation} X(e_ir+e_j,\,e_k)=X(e_i,\,rae_k)X(e_j,\,ae_k)\,\text{ for }r\in R,\ a\in I\text{ and } i,j,k\text{ distinct}. \label{add3'} \end{equation} \end{rk}

@@

\section{The local-global principle for $\Kt$.} \label{sec:lgp}
%At this point we have demonstrated for $n\geq4$ and a splitting ideal $I$ that $\St^*(n,\,R,\,I)$ and $\St(n,\,R,\,I)$ are isomorphic.
Now we turn to the main result of this paper, namely, we construct the map $$T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X]),\ \text{ for} n\geq4.$$ 
The local-global principle for the Steinberg group and centrality of $\mathrm K_2$ will formally follow from the existence of this map, see~\cite[Lemmas~15--16, and the proof of Theorem~2]{SCh}.

We will prove a somewhat more general result. Let $B$ be a commutative unital ring.
A $B$-module $M$ is called \emph{$r$-divisible} if for every $m\in M$ there exists $m'\in M$ such that $rm' = m$ (we denote such $m'$ by $\frac{m}{a}$).
Clearly, $M$ is $a$-divisible if and only if the morphism of principal localization $\lambda_a\colon M \to M_a$ is an isomorphism.

\begin{tm} \label{a3map} Let $B$ be a ring, $a\in B$ and let $I$ be an ideal $B$ that is $a$-divisible.
Then there exists a map $T\colon\St(n,\,B_a,\,I)\rightarrow\St(n,\,B)$ which makes the following diagram commute.
$$\xymatrix{ \St^*(2n,\,B,\,I)\ar@<-0.0ex>[rr]^{\iota}\ar@<-0.0ex>[d]_{\lambda_{a}^*}&&\St(2n,\,B)\ar@<-0.0ex>[d]^{\lambda_{a}^*}\\
             \St^*(2n,\,B_a,\,I)\ar@<-0.0ex>[rr]_{\iota}\ar@{-->}[rru]_{T}&&\St(2n,\,B_a)}$$ \end{tm}

To prove this theorem we need to find some elements ``$X(u,\,v)$'' inside $\St(n,\,B)$ for $u\in\E(n,\,B_a)e_1$, $v\in I$. The idea is the following. Say $a$ is not a zero divisor, then $ua^m\in B$ for some $m$ and we will construct some element ``$X(ua^m,\,va^{-m})$''. One can not garantee that $ua^m\in\E(n,\,B)e_1a^k$, however, the ideal generated by entries of $ua^m$ contains some power of $a$ (this is equivalent to say that $ua^m$ becomes unimodular after the principle localisation in $a$). Denote $\mathrm I(u)$ the ideal generated by entries of $u\in R^n$, $\mathrm I(u)=\sum\limits_{k=1}^nu_kR$. Then, we need generators parametrised by pairs $(u,\,v)$ with $u^tv=0$, $v\in I$, $a^m\in\mathrm I(u)$ for some $m\in\mathbb N$. Such generators are defined in~\cite{Tul}.

\begin{df}[Tulenbaev]
For $u$, $v\in B^n$, $a\in\mathrm I(u)$ and $v_1,\ldots,v_N\in B^n$ such that $u^tv_k=0$ $\forall\,k$,  each $v_k$ has at least two zero coordinates and $\sum_{k=1}^Nv_k=v$ define $X_{u,v}(a)=\prod\limits_{k=1}^Nx(u,\,v_ka)$. Tulenbaev shows that factors $x(u,\,v_ka)$ commute~\cite[Lem.~1.1\,e)]{Tul} and that for any other decomposition of $v$ as a sum $v=\sum_{j=1}^Mv'_j$ of vectors orthogonal to $u$ and having two zero coordinates holds $\prod_{k=1}^Nx(u,\,v_ka)=\prod_{j=1}^Mx(u,\,v'_ja)$~\cite[p.~3]{Tul}, i.e., elements $X_{u,v}(a)$ are well-defined. Obviously, $\phi(X_{u,v}(a))=t(u,\,va)$.
\end{df}

Observe that Tulenbaev uses different notation for van der Kallen elements. He writes $X_{u,v}$ instead of $X(u,\,v)$ and $X(u,\,v)$ instead of $x(u,\,v)$. We keep van der Kallen's notation.

\begin{lm}
\label{xproperties}
For $u$, $v$, $v'$ and $w\in B^n$, such that $v$ and $v'$ have decomposition as in above definition, $u^tw=0$, $a$, $b\in\mathrm I(u)$, $c\in B$, $g\in\St(n,\,B)$ holds
\begin{enumerate}
\item
$X_{u,vc}(a)=X_{u,v}(ca)$,
\item
$X_{uc,v}(ca)=X_{u,vc^2}(a)$,
\item
$X_{u,v}(a)X_{u,v'}(a)=X_{u,v+v'}(a)$,
\item
$g\,X_{u,wb}(a)g\inv=X_{\phi(g)u,\phi(g)^*wb}(a)$.
\end{enumerate}
\end{lm}

>>>>

\begin{rk} \label{rk:T123'} Notice that from the proof of the above proposition it follows that one could replace the family of relations \eqref{add3}
in the definition of $\St(n, R, I)$ with the following smaller family:
\setcounter{equation}{2} \renewcommand{\theequation}{T\arabic{equation}'}
\begin{equation} X(e_ir+e_j,\,e_k)=X(e_i,\,rae_k)X(e_j,\,ae_k)\,\text{ for }r\in R,\ a\in I\text{ and } i,j,k\text{ distinct}. \label{add3'} \end{equation} \end{rk}

@@

\section{The local-global principle for $\Kt$.} \label{sec:lgp}
%At this point we have demonstrated for $n\geq4$ and a splitting ideal $I$ that $\St^*(n,\,R,\,I)$ and $\St(n,\,R,\,I)$ are isomorphic.
Now we turn to the main result of this paper, namely, we construct for $n\geq 4$ the map $$T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X]),$$ that fits into the diagram \eqref{eq:diagram}.
The local-global principle for the Steinberg group and centrality of $\mathrm K_2$ will formally follow from the existence of this map, see~\cite[Lemmas~15--16, and the proof of Theorem~2]{SCh}.

We will prove a somewhat more general result. Let $B$ be a commutative unital ring.
A $B$-module $M$ is called \emph{$r$-divisible} if for every $m\in M$ there exists $m'\in M$ such that $rm' = m$ (we denote such $m'$ by $\frac{m}{a}$).
Clearly, $M$ is $r$-divisible if and only if the morphism of principal localisation $\lambda_r\colon M \to M_r$ is an isomorphism.

\begin{tm} \label{a3map} Let $B$ be a ring, $a\in B$ and let $I$ be an ideal $B$ that is $a$-divisible.
Then for $n\geq 4$ there exists a map $T\colon\St(n,\,B_a,\,I)\rightarrow\St(n,\,B)$ which makes the following diagram commute.
$$\xymatrix{ \St^*(n,\,B,\,I)\ar@<-0.0ex>[rr]^{\iota}\ar@<-0.0ex>[d]_{\lambda_{a}^*} && \St(n,\,B)\ar@<-0.0ex>[d]^{\lambda_{a}^*}\\
             \St^*(n,\,B_a,\,I)\ar@<-0.0ex>[rr]_{\iota}\ar@{-->}[rru]_{T}            && \St(n,\,B_a)}$$ \end{tm}

For a vector $u\in R^n$ we denote by $I(u)$ the ideal generated by the entries of $u\in R^n$, i.\,e. $I(u)=\sum\limits_{k=1}^nu_kR$.
To prove \cref{a3map} we follow the approach of Tulenbaev and construct yet another family of elements $X_{u,v}(a)\in \St(n,\,B)$, $a\in B$ which contains van der Kallen's generators $X(u, v)$ as a subfamily (i.\,e. $X_{u,v}(1) = X(u, v)$).
The key feature of this new family is that the assumption $u \in\Um(n, B)$ is replaced with a weaker condition $a \in I(u)$ (this is equivalent to saying that $u$ becomes unimodular after localisation in $a$).
%The idea is the following. Say $a$ is not a zero divisor, then $ua^m\in B$ for some $m$ and we will construct some element ``$X(ua^m,\,va^{-m})$''.
%One can not garantee that $ua^m\in\E(n,\,B)e_1a^k$, however, the ideal generated by entries of $ua^m$ contains some power of $a$ .
More formally, we have the following definition.
%Then, we need generators parametrised by pairs $(u,\,v)$ with $u^tv=0$, $v\in I$, $a^m\in\mathrm I(u)$ for some $m\in\mathbb N$. Such generators are defined in~\cite{Tul}.

\begin{df}[Tulenbaev]
Let $u$, $v\in B^n$ be such that $v=\sum_{k=1}^Nv_k$ for some $v_1,\ldots,v_N\in B^n$ such that $u^tv_k=0$ and $v_k$ has at least two zero coordinates for all $k$.
We call such a decomposition of $v$ a \emph{canonical decomposition}.
Now for $a\in I(u)$ define $X_{u,v}(a) = \prod\limits_{k=1}^Nx(u,\,v_ka)$. 
\end{df}
Tulenbaev shows that the factors $x(u,\,v_ka)$ commute~\cite[Lemma~1.1\,e)]{Tul} and that for any other canonical decomposition $v=\sum_{j=1}^Mv'_j$ one has
$\prod_{k=1}^Nx(u,\,v_ka)=\prod_{j=1}^Mx(u,\,v'_ja)$, see the discussion after \cite[Lemma~1.1]{Tul}.
Thus, the elements $X_{u,v}(a)$ are well-defined. Obviously, $\phi(X_{u,v}(a))=t(u,\,va)$.

Notice that Tulenbaev uses different notation for van der Kallen elements, e.\,g. he writes $X_{u,v}$ instead of $X(u,\,v)$ and $X(u,\,v)$ instead of $x(u,\,v)$. 
We stick to van der Kallen's notation. \textbf{TODO: I thought we use our own notation?}

\begin{lm}
\label{xproperties}
For $u$, $v$, $v'$ and $w\in B^n$, such that $v$ and $v'$ possess a canonical decomposition, $u^tw=0$, $a$, $b\in I(u)$, $c\in B$, $g\in\St(n,\,B)$ the following facts are true.
\begin{enumerate}
\item $X_{u,vc}(a)=X_{u,v}(ca)$,
\item $X_{uc,v}(ca)=X_{u,vc^2}(a)$,
\item $X_{u,v}(a)X_{u,v'}(a)=X_{u,v+v'}(a)$,
\item $g\,X_{u,wb}(a)g\inv=X_{\phi(g)u,\phi(g)^*wb}(a)$.
\end{enumerate}
\end{lm}
